---
title: "argos-blvu - 4. data analysis BLVU AGL analysis "
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


clear environment 
```{r}
rm(list = ls())

```


libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
library(snow)
```


read the data for analysis 
```{r}
tic()

blvu1<- fread("blvu1.csv")
colnames(blvu1)

cat("
Ran at: ");Sys.time()
toc()
```
Identify flying and perch locations
p = perch f = fly n = not used
```{r}
tic()

blvu1$p <- if_else((blvu1$kph < 1 & blvu1$agl < 50),"p", 
                   if_else((blvu1$agl < -50), "n", "f"))

fly1<- subset(blvu1, blvu1$p == "f")


summary(fly1)

cat("
Ran at: ");Sys.time()
toc()

```

Add squared terms to dataset


```{r}
tic()

fly1$dem2 <- fly1$dem^2
fly1$slope2 <- fly1$slope^2
fly1$north2 <- fly1$north^2
fly1$east2 <- fly1$east^2
fly1$drd2 <- fly1$drd^2
#center and scale squared terms
fly1$dem2s <- scale(fly1$dem2, center = T, scale = T)
fly1$slope2s <- scale(fly1$slope2, center = T, scale = T)
fly1$drd2s <- scale(fly1$drd2, center = T, scale = T)


cat("
Ran at: ");Sys.time()
toc()
```




Look at agl data 
and transformed agl
```{r}
tic()

ggplot(fly1, aes(agl)) + geom_boxplot()

cat("
Ran at: ");Sys.time()
toc()
```

Analysis of data for BLVU AGL

 
 
```{r}
tic()

glmm1 <- lme4::lmer (agl ~ dems + dem2s + east + east2 + north + north2 + drds + drd2s + nlcdcat + tpicat + tricat + (1|id) , data = fly1 , na.action = na.fail)

summary(glmm1)

cat("
Ran at: ");Sys.time()
toc()
```

check fit of model

```{r}
tic()

#residuals
rglmm1<- residuals (glmm1, type = "pearson")
#qqplot of residuals
qqnorm (rglmm1)
qqline (rglmm1)
# predicted fit
fglmm1 <- fitted(glmm1)
#plot residuals vs predicted fit
plot(fglmm1,rglmm1)


cat("
Ran at: ");Sys.time()
toc()
```

dredge pdredge for mulit-core analysis


```{r}
tic()
#set up the cluster
cores<-6
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
cluster <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
#export data to cluster
clusterExport(cluster,"fly1")
#explort package to cluster
clusterEvalQ(cluster, library(lme4))

#run dredge
fly.dredge <- pdredge(glmm1, rank = "AICc", cluster=cluster)
# fly.dredge <- dredge(glmm1, rank = "AICc")

#stop cluster
stopCluster(cluster)

#write results
fwrite(fly.dredge, "fly.dredge.csv")

cat("
Ran at: ");Sys.time()
toc()
```


model average


```{r}
tic()

mafly <- model.avg(fly.dredge)

cat("
Ran at: ");Sys.time()
toc()
```

Look at model averaged results 

```{r}
tic()

summary(mafly)

cat("
Ran at: ");Sys.time()
toc()
```