---
title: "argos-blvu - 3. data analysis BLVU RSF "
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


clear environment 
```{r}
rm(list = ls())

```


libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
library(snow)
```


read the data for analysis 
```{r}
tic()

data1<- fread("data1_1.csv")


cat("
Ran at: ");Sys.time()
toc()
```


Analysis of data for BLVU RSF


```{r}
tic()

glm1 <- glm( y ~ -1 +  slopes + flat*east + flat*north + dems + tpicat + tricat + nlcdcat + dprds + dsrds + dforests + dsolids , family = binomial  , data = data1, na.action = na.fail)

summary(glm1)



cat("
Ran at: ");Sys.time()
toc()
```

check fit of model

```{r}
tic()

#residuals
rglm1<- residuals (glm1, type = "pearson")
#qqplot of residuals
qqnorm (rglm1)
qqline (rglm1)
# predicted fit
fglm1 <- fitted(glm1)
#plot residuals vs predicted fit
plot(fglm1,rglm1)


cat("
Ran at: ");Sys.time()
toc()
```

model 2
 ```{r}
# tic()
# 
# glm2 <- glm( y ~ -1 +  slopes + east + north + dems + tpicat + tricat + nlcdcat + drds , family = binomial  , data = data1, na.action = na.fail)
# 
# summary(glm2)
# 
# 
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```
check fit of model

```{r}
tic()

# #residuals
# rglm2<- residuals (glm2, type = "pearson")
# #qqplot of residuals
# qqnorm (rglm2)
# qqline (rglm2)
# # predicted fit
# fglm2 <- fitted(glm2)
# #plot residuals vs predicted fit
# plot(fglm2,rglm2)
# 
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


dredge results


```{r}
tic()

#set up the cluster
cores<-6
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
cluster <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
#export data to cluster
clusterExport(cluster,"data1")
#explort package to cluster
clusterEvalQ(cluster, library(stats))

#run dredge
data.dredge <- pdredge(glm1, rank = "AICc", cluster=cluster)


#stop cluster
stopCluster(cluster)

#write results
fwrite(data.dredge, "data.dredge.csv")



cat("
Ran at: ");Sys.time()
toc()
```


model average

```{r}
tic()


madredge1 <- model.avg(data.dredge)


cat("
Ran at: ");Sys.time()
toc()
```


View results of dredge


```{r}
tic()

summary(data.dredge)
summary(madredge1)

cat("
Ran at: ");Sys.time()
toc()
```





```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```





```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```





```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```



