---
title: "argos-blvu - 1. Data setup"
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
```

import data
```{r}
tic()

#import
blvu1<- fread("wv_blvu.csv")
random1<- fread("random.csv")


cat("
Ran at: ");Sys.time()
toc()
```


save original data

```{r}
tic()

blvu.orig <- blvu1
random.orig <- random1


cat("
Ran at: ");Sys.time()
toc()
```

Filter data 
Remove poor quality points
HDOP&VDOP >10
2d points
```{r}
tic()

blvu1a<- subset(blvu1, blvu1$Fix>2)
blvu1b <- subset(blvu1a, blvu1a$HDOP <= 10)
blvu1c <- subset(blvu1b, blvu1b$VDOP <= 10)
nrow(blvu1)
nrow(blvu1a)
nrow(blvu1b)
nrow(blvu1c)

blvu1 <- blvu1c


#cleanup 
rm(list = ls()[! ls() %in% c("blvu1","blvu.orig", "random1", "random.orig")])


cat("
Ran at: ");Sys.time()
toc()
```


setup blvu data
```{r}
tic()

colnames(blvu1)

blvu1$y <- 1

names(blvu1)[names(blvu1)=="Aspect_12"]<-"aspect"
names(blvu1)[names(blvu1)=="DEM_12"]<-"dem"
names(blvu1)[names(blvu1)=="SLOPE_12"]<-"slope"
names(blvu1)[names(blvu1)=="TPI_12"]<- "tpi"
names(blvu1)[names(blvu1)=="TRI_12"]<-"tri"
names(blvu1)[names(blvu1)=="EucDist_PriS1_12"]<-"drd"
names(blvu1)[names(blvu1)=="nlcd_2019_land_cover_l48_20210604_25517"]<-"nlcd"
names(blvu1)[names(blvu1)=="Month_LocalTime"]<-"month"
colnames(blvu1)

#add season variable 

blvu1$season <- if_else(blvu1$month == 12, "w", 
                if_else(blvu1$month == 1, "w",
                if_else(blvu1$month == 2,"w",
                if_else(blvu1$month == 3, "s",
                if_else(blvu1$month == 4, "s",
                if_else(blvu1$month ==  5, "s",
                if_else(blvu1$month ==  6, "m",
                if_else(blvu1$month == 7, "m",
                if_else(blvu1$month == 8, "m", "f"          
                          )))))))))


cat("
Ran at: ");Sys.time()
toc()
```

get random points set up correctly with the same variables as blvu data
combine the two datasets
```{r}
tic()

#rename columns in blvu data
blvu2 <- blvu1 %>% select (id = Animal_ID,
                           dt = Date_Time,
                           lat = Latitude,
                           lon = Longitude,
                           alt = Altitude,
                           aspect = aspect,
                           tri = tri,
                           dem = dem,
                           tpi = tpi,
                           slope = slope,
                           drd = drd,
                           nlcd = nlcd,
                           kph = KPH,
                           date = Date,
                           season = season,
                           y = y,
                           )

colnames(random1)
# retain only columns that we need 
rand2 <- random1 %>% select (lat = Lat,
                           lon = Lon,
                           aspect = Aspect,
                           tri = TRI,
                           dem = DEM,
                           tpi = TPI,
                           slope = SLOPE,
                           drd = drd,
                           nlcd = nlcd,
                           )


# add additional variables needed
colnames(blvu2)
blvu2$y <- 1
colnames(rand2)
rand2$id <- 999
rand2$dt <- 9/9/9999
rand2$kph <- 0 
rand2$alt <- 0
rand2$y <- 0 
rand2$date <- 0
rand2$season <- "n"


data2 <- rbind(rand2, blvu2)

random1<- rand2
blvu1 <- blvu2

#remove nas from data
data1 <- na.omit(data2)
#cleanup 
rm(list = ls()[! ls() %in% c("blvu1","blvu.orig", "random1", "random.orig", "data1")])



cat("
Ran at: ");Sys.time()
toc()
```

Put together additional variables needed for analysis
topography


```{r}
tic()

data1a <- subset(data1, aspect>=0)
data1 <- data1a
rm(data1a)

 data1$north <- cos(data1$aspect*pi/180)
 data1$east <- sin(data1$aspect*pi/180)




data1$nlcdcat <- as.factor(data1$nlcd)
summary(data1$nlcdcat)
data1$nlcdcat <- as.factor(if_else(data1$nlcd ==11, "water",
if_else(data1$nlcd ==21, "devel",
if_else(data1$nlcd ==22, "devel",
if_else(data1$nlcd ==23, "devel",
if_else(data1$nlcd ==24, "devel",
if_else(data1$nlcd ==31, "bare",
if_else(data1$nlcd ==41, "forest",
if_else(data1$nlcd ==42, "forest",
if_else(data1$nlcd ==43, "forest",
if_else(data1$nlcd ==52, "shrub",
if_else(data1$nlcd ==71, "grass",
if_else(data1$nlcd ==81, "ag", 
if_else(data1$nlcd ==82, "ag",         
if_else(data1$nlcd ==90, "wetland",         
if_else(data1$nlcd ==95, "wetland", 
        
        "PROBLEM NLCD"))))))))))))))))
summary(data1$nlcdcat)


data1$smootr <- if_else(data1$tri <= 80, 1, 0)
data1$slighr <- if_else(data1$tri > 80 & data1$tri <= 160, 1, 0)
data1$lowr <- if_else(data1$tri > 160 & data1$tri <= 240, 1, 0)
data1$moderr <- if_else(data1$tri > 240 & data1$tri <= 500, 1, 0)
data1$highr <- if_else(data1$tri > 500, 1, 0)

data1$tricat <- as.factor(if_else(data1$smootr == 1, "smooth", 
                                 if_else(data1$slighr == 1, "slight",
                                         if_else(data1$lowr == 1, "low", 
                                                 if_else(data1$moderr == 1, "moder", 
                                                         if_else(data1$highr == 1, "high","PROBLEM tri"))))))

summary(data1$tricat)


data1$valley <- if_else(data1$tpi < -1,1,0)
data1$ridge <- if_else(data1$tpi > 1 ,1,0)
data1$gentle <- if_else(data1$tpi >= -1 & data1$tpi <= 1 & data1$slope<=0.06 ,1,0)
data1$steep <- if_else(data1$tpi >= -1 & data1$tpi <= 1 & data1$slope>0.06,1,0)

data1$tpicat <- as.factor(if_else(data1$valley == 1, "valley",
                                 if_else(data1$ridge == 1, "ridge",
                                         if_else(data1$gentle== 1, "gentle",
                                                 if_else(data1$steep== 1, "steep","PROBLEM tpi")))))

summary(data1$tpicat)


data1$agl <- data1$alt - data1$dem


cat("
Ran at: ");Sys.time()
toc()
```

Save data for analysis later

```{r}
tic()
blvu1 <- subset(data1, data1$y == 1)
random1 <- subset(data1, data1$y ==0)

fwrite(blvu1, "blvu1.csv")
fwrite(random1, "random1.csv")
fwrite(data1, "data1.csv")

cat("
Ran at: ");Sys.time()
toc()
```