---
title: "argos-blvu - 1. Vulture data checking"
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
```

import data
```{r}
tic()

#import
blvu1<- fread("wv_blvu.csv")
blvu2<- fread("new_blvu.csv")

cat("
Ran at: ");Sys.time()
toc()
```


Filter data 
Remove poor quality points
HDOP&VDOP >10
2d points
```{r}
tic()

blvu1a<- subset(blvu1, blvu1$Fix>2)
blvu1b <- subset(blvu1a, blvu1a$HDOP <= 10)
blvu1c <- subset(blvu1b, blvu1b$VDOP <= 10)
nrow(blvu1)
nrow(blvu1a)
nrow(blvu1b)
nrow(blvu1c)

blvu1 <- blvu1c

blvu2a<- subset(blvu2, blvu2$Fix>2)
blvu2b <- subset(blvu2a, blvu2a$HDOP <= 20)
blvu2c <- subset(blvu2b, blvu2b$VDOP <= 20)
nrow(blvu2)
nrow(blvu2a)
nrow(blvu2b)
nrow(blvu2c)

blvu2 <- blvu2c




cat("
Ran at: ");Sys.time()
toc()
```

view data to see if there are outliers

```{r}
tic()

blvu1$altdiffrate <- blvu1$AltDiff/blvu1$TimeDiff
blvu1$mdiffrate <- blvu1$M_fromLast/blvu1$TimeDiff


ggplot(blvu1, aes(altdiffrate)) +
  geom_boxplot()

ggplot(blvu1, aes(mdiffrate)) +
  geom_boxplot()

blvu2$altdiffrate <- blvu2$AltDiff/blvu2$TimeDiff
blvu2$mdiffrate <- blvu2$M_fromLast/blvu2$TimeDiff


ggplot(blvu2, aes(altdiffrate)) +
  geom_boxplot()

ggplot(blvu2, aes(mdiffrate)) +
  geom_boxplot()


cat("
Ran at: ");Sys.time()
toc()
```
write files to look at in GIS

```{r}
tic()

fwrite(blvu1, "blvu1.intermediate.csv")
fwrite(blvu2, "blvu2.intermediate.csv")

cat("
Ran at: ");Sys.time()
toc()
```

remove observations with very high or low altdiffrates > |40|
```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```


```{r}
tic()

cat("
Ran at: ");Sys.time()
toc()
```





Save data for analysis later

```{r}
tic()

fwrite(blvu1, "blvu1checked.csv")
fwrite(blvu2, "blvu2checked.csv")


cat("
Ran at: ");Sys.time()
toc()
```