---
title: "argos-blvu - 2. data exploration "
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


clear environment 
```{r}
rm(list = ls())

```


libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
```

Load the data
```{r}
tic()

data1<- fread("data1_1.csv")
blvu1 <- fread("blvu1_1.csv")

cat("
Ran at: ");Sys.time()
toc()
```


Setup theme for ggplot

```{r}
tic()

thm <- theme_classic(base_size = 14, base_family = "serif") + 
  theme(strip.background = element_blank()) + 
  theme(axis.text = element_text( size = 10 ), 
        axis.title = element_text( size = 12),
        strip.text = element_text(size = 12))

dev.new(width = 7, height = 2.5, unit = "in")

cat("
Ran at: ");Sys.time()
toc()
```



examine data to look at relationships among variables

```{r}
tic()


a<- ggplot(data1, aes(as.factor(y), north))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), east))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), slope))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), dem))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), dsrd))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), dprd))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), dforest))
a + geom_violin()

a<- ggplot(data1, aes(as.factor(y), dsolid))
a + geom_violin()

a<- ggplot(blvu1, aes(agl))
a + geom_histogram() 



cor <- cor(data1[,c("lat","lon", "north", "east", "dem", "tri", "tpi", "slope", "dsrd", "dprd", "dforest", "dsolid")])

cor


cat("
Ran at: ");Sys.time()
toc()
```


more summaries 
```{r}
tic()

a<- ggplot(data1, aes(tricat)) 
a + geom_bar()+ facet_grid(~y)

a<- ggplot(data1, aes(tpicat))
a + geom_bar()+ facet_grid(~y)

a<- ggplot(data1, aes(nlcdcat))
a + geom_bar()+ facet_grid(~y)

cat("
Ran at: ");Sys.time()
toc()
```







export correlation matrix
```{r}
tic()

fwrite(cor, "cor.csv")

cat("
Ran at: ");Sys.time()
toc()
```



