---
title: "argos-blvu - 1. Vulture data checking"
author: "Adam Duerr"
date: "`r Sys.Date()`"
output: html_document
---

notes: try to use all lowercase for files just to simplify typing and retyping the names
use git and renv to track changes to the project
git repo is https:/github.com/adamdaz/argos-blvu/

add info to each chunk to track whether it was run recently and how long it took
```{r}
# tic()
# 
# cat("
# Ran at: ");Sys.time()
# toc()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


clear environment 
```{r}
rm(list = ls())

```


libraries
```{r}
library(usethis) #git stuff
library(renv) #versioning tools
library(tictoc) #tracking time
library(data.table) #allows multi-core processing for reading/writing tables = fast!
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ResourceSelection)
library(MuMIn)
```

import data
```{r}
tic()

#import
blvu1<- fread("blvu.csv")
blvu2<- fread("new_blvu.csv")

colnames(blvu1)
colnames(blvu2)

cat("
Ran at: ");Sys.time()
toc()
```


Filter data 
Remove poor quality points
HDOP&VDOP >10
2d points
high altitude >2500 but low speed points <5 kph
very high altitude >5000
very fast speeds >125 kph
BuckLee (animal id = 37) after 1/4/2020
```{r}
tic()



blvu1a<- subset(blvu1, blvu1$Fix>2)
blvu1b <- subset(blvu1a, blvu1a$HDOP <= 10)
blvu1c <- subset(blvu1b, blvu1b$VDOP <= 10)
nrow(blvu1)
nrow(blvu1a)
nrow(blvu1b)
nrow(blvu1c)

blvu1d <- blvu1c[!(blvu1c$Altitude > 2500 & blvu1c$KPH < 5)]
blvu1e <- blvu1d[!(blvu1d$Altitude > 5000|blvu1d$KPH > 125)]
blvu1f <- blvu1e[!(blvu1e$Animal_ID == 37 & blvu1e$Date>01/04/2020 )]
nrow(blvu1e)
nrow(blvu1f)
blvu1 <- blvu1f

blvu2a<- subset(blvu2, blvu2$Fix>2)
blvu2b <- subset(blvu2a, blvu2a$HDOP <= 20)
blvu2c <- subset(blvu2b, blvu2b$VDOP <= 20)
nrow(blvu2)
nrow(blvu2a)
nrow(blvu2b)
nrow(blvu2c)
blvu2d <- blvu2c[!(blvu2c$Altitude > 2500 & blvu2c$KPH < 5)]
blvu2e <- blvu2d[!(blvu2d$Altitude > 5000|blvu2d$KPH > 125)]
blvu2f <- blvu2e[!(blvu2e$Animal_ID == 37 & blvu2e$Date>01/04/2020 )]
nrow(blvu2e)
nrow(blvu2f)
blvu2 <- blvu2f




cat("
Ran at: ");Sys.time()
toc()
```

view data to see if there are outliers

```{r}
tic()

blvu1$altdiffrate <- blvu1$AltDiff/blvu1$TimeDiff
blvu1$mdiffrate <- blvu1$M_fromLast/blvu1$TimeDiff


ggplot(blvu1, aes(altdiffrate)) +   geom_boxplot()

ggplot(blvu1, aes(mdiffrate)) +   geom_boxplot()

ggplot(blvu1, aes(Altitude)) + geom_boxplot()

ggplot(blvu1, aes(KPH)) + geom_boxplot()



cat("
Ran at: ");Sys.time()
toc()
```



```{r}
tic()


blvu2$altdiffrate <- blvu2$AltDiff/blvu2$TimeDiff
blvu2$mdiffrate <- blvu2$M_fromLast/blvu2$TimeDiff


ggplot(blvu2, aes(altdiffrate)) +   geom_boxplot()

ggplot(blvu2, aes(mdiffrate)) +   geom_boxplot()

ggplot(blvu2, aes(Altitude)) +   geom_boxplot()

ggplot(blvu2, aes(KPH)) +   geom_boxplot()


cat("
Ran at: ");Sys.time()
toc()
```




write files to look at in GIS

```{r}
tic()

fwrite(blvu1, "blvu1.intermediate.csv")
fwrite(blvu2, "blvu2.intermediate.csv")

cat("
Ran at: ");Sys.time()
toc()
```

```{r}
tic()



cat("
Ran at: ");Sys.time()
toc()
```




Save data for analysis later

```{r}
tic()

fwrite(blvu1, "blvu1_0.csv")
fwrite(blvu2, "blvu2_0.csv")


cat("
Ran at: ");Sys.time()
toc()
```